TITLE

### 2024-04-26 papers/Draft/draft.tex
Adding tables to the paper
Added values to the tables

%%BIBLIOGRAPHY- This uses biber/biblatex to generate bibliographies according to the
%%Unified Style Sheet for Linguistics
\usepackage[main=american, german]{babel}% Recommended
\usepackage{csquotes}% Recommended
\usepackage[backend=biber,
        style=unified,
        maxcitenames=3,
        maxbibnames=99,
        natbib,
        url=false]{biblatex}
\addbibresource{Dissertation.bib}
\setcounter{biburlnumpenalty}{100}  % allow URL breaks at numbers
% \setcounter{biburlucpenalty}{100}   % allow URL breaks at uppercase letters
% \setcounter{biburllcpenalty}{100}   % allow URL breaks at lowercase letters

@@

Santiago Laxopa Zapotec is an Oto-manguean

@@

If the absolute value of f0 was greater than 3, it was removed. This is because 99.7\% of the data in a normally distributed dataset lies within 3 SDs of the mean. Anything greater than 3 is likely an outlier and marked as such.
Formants
Data was again grouped by speaker, and Mahalanobis distance was calculated for F1 and F2. 
A Mahalanobis distance greater than 6 means that you are a likely outlierxc
This was done by taking the covariance and means of F1 and F2. This gives you a grouping based on the vowels’ formants. 
The Mahalanobis distance was calculated based on F1 and F2
The data was filtered by each vowel and then outliers were determined. 
Energy
If energy was equal to 0 it was converted to NA
I then took the log10  of energy across all datapoints because the data is left bounded by 0 and has a long right tail. 
After determining which items were outliers they were filtered out.

@@

\begin{table}[!h]
  \centering
  \caption{Results of the statistical model for H1*-H2*.}
  \label{tab:H1H2Results}
  \begin{tabular}{llllll}
    \lsptoprule
     & Estimate & Std. Error & df & t value & p-value \\ \hline
    \textit{Intercept} & 0.02819 & 0.1033 & 4.755 & 0.273 & 0.796 \\
    Breathy &&&&& \\
    Checked &&&&& \\
    Laryngealized &&&&& \\
    Position 2 &&&&& \\
    Position 3 &&&&& \\
    High tone &&&&& \\
    Low tone &&&&& \\
    Mid tone &&&&& \\
    Rising tone &&&&& \\
    Breathy:Position 2 &&&&& \\
    Checked:Position 2 &&&&& \\
    Laryngealized:Position 2 &&&&& \\
    Breathy:Position 3 &&&&& \\
    Checked:Position 3 &&&&& \\
    Laryngealized:Position 3 &&&&& \\
    \lspbottomrule
  \end{tabular}
\end{table}

@@

\begin{table}[!h]
  \centering
  \caption{Results of the statistical model for H1*-A3.}
  \label{tab:H1A3Results}
  \begin{tabular}{llllll}
    \lsptoprule
     & Estimate & Std. Error & df & t value & p-value \\ \hline
    \textit{Intercept} & 0.02819 & 0.1033 & 4.755 & 0.273 & 0.796 \\
    Breathy &&&&& \\
    Checked &&&&& \\
    Laryngealized &&&&& \\
    Position 2 &&&&& \\
    Position 3 &&&&& \\
    High tone &&&&& \\
    Low tone &&&&& \\
    Mid tone &&&&& \\
    Rising tone &&&&& \\
    Breathy:Position 2 &&&&& \\
    Checked:Position 2 &&&&& \\
    Laryngealized:Position 2 &&&&& \\
    Breathy:Position 3 &&&&& \\
    Checked:Position 3 &&&&& \\
    Laryngealized:Position 3 &&&&& \\
    \lspbottomrule
  \end{tabular}
\end{table}

@@

\begin{table}[!h]
  \centering
  \caption{Results of the statistical model for Residual H1*.}
  \label{tab:H1H2Results}
  \begin{tabular}{llllll}
    \lsptoprule
     & Estimate & Std. Error & df & t value & p-value \\ \hline
    \textit{Intercept} & 0.02819 & 0.1033 & 4.755 & 0.273 & 0.796 \\
    Breathy &&&&& \\
    Checked &&&&& \\
    Laryngealized &&&&& \\
    Position 2 &&&&& \\
    Position 3 &&&&& \\
    High tone &&&&& \\
    Low tone &&&&& \\
    Mid tone &&&&& \\
    Rising tone &&&&& \\
    Breathy:Position 2 &&&&& \\
    Checked:Position 2 &&&&& \\
    Laryngealized:Position 2 &&&&& \\
    Breathy:Position 3 &&&&& \\
    Checked:Position 3 &&&&& \\
    Laryngealized:Position 3 &&&&& \\
    \lspbottomrule
  \end{tabular}
\end{table}

>>>>

%%BIBLIOGRAPHY- This uses biber/biblatex to generate bibliographies according to the
%%Unified Style Sheet for Linguistics
\usepackage[main=american, german]{babel}% Recommended
\usepackage{csquotes}% Recommended
\usepackage[backend=biber,
        style=apa,
        maxcitenames=3,
        maxbibnames=99,
        natbib,
        url=false]{biblatex}
\addbibresource{Dissertation.bib}
\setcounter{biburlnumpenalty}{100}  % allow URL breaks at numbers
% \setcounter{biburlucpenalty}{100}   % allow URL breaks at uppercase letters
% \setcounter{biburllcpenalty}{100}   % allow URL breaks at lowercase letters

@@

Santiago Laxopa Zapotec is a Northern Zapotec language of Oto-manguean language family \citep{adlerAcousticsPhonationTypes2016,adlerDerivationVerbInitiality2018,foleyForbiddenCliticClusters2018,foleyExtendingPersonCaseConstraint2020,sichelPronounsAttractionSierra2020,
sichelFeaturalLifeNominals2020,brinkerhoffDownstepSantiagoLaxopaMFM,brinkerhoffTonalPatternsTheir2022}. It is spoken by 981 people in the municipality of Santiago Laxopa, Ixtlán, Oaxaca, Mexico \citep{SantiagoLaxopaEconomya} and a small number of other speakers in diaspora throughout Mexico and the United States. Similar to other Oto-manguean languages, Santiago Laxopa Zapotec is laryngeally complex. According to \citet{silvermanLaryngealComplexityOtomanguean1997,silvermanPhasingRecoverability1997,blankenshipTimeCourseBreathiness1997,blankenshipTimingNonmodalPhonation2002}, laryngeal complexity refers to how these languages make use of both contrastive tone and contrastive voice quality.

@@

If the absolute value of f0 was greater than 3, it was removed. This is because 99.7\% of the data in a normally distributed dataset lies within 3 SDs of the mean. Anything greater than 3 is likely an outlier and marked as such.
Formants
Data was again grouped by speaker, and the Mahalanobis distance \citep{martosGeneralizationMahalanobisDistance2013,
mahalanobisGeneralizedDistanceStatistics2018,drumondUsingMahalanobisDistance2019} was calculated for F1 and F2. 
A Mahalanobis distance greater than 6 means that you are a likely outlier
This was done by taking the covariance and means of F1 and F2. This gives you a grouping based on the vowels’ formants. 
The Mahalanobis distance was calculated based on F1 and F2
The data was filtered by each vowel and then outliers were determined. 
Energy
If energy was equal to 0 it was converted to NA
I then took the log10  of energy across all datapoints because the data is left bounded by 0 and has a long right tail. 
After determining which items were outliers they were filtered out.

@@

\begin{table}[!h]
  \centering
  \caption{Results of the statistical model for H1*-H2*.}
  \label{tab:H1H2Results}
  \begin{tabular}{lllllll}
    \lsptoprule
     & Estimate & Std. Error & df & t value & p-value & \\ \hline
    \textit{Intercept} & 0.02819 & 0.1033 & 4.755 & 0.273 & 0.796 & \\
    Breathy & -3.446e-02 & 4.21e-02&9.821e+03&-0.818& 0.413507 & \\
    Checked & -1.402e-01 & 4.055e-02 & 1.465e+04 & -3.457 & 0.000547 & *** \\
    Laryngealized & -1.361e-01 &4.969e-02 & 5.348e+03 & -2.740 & 0.006174 & ** \\
    Position 2 & 4.749e-03 & 1.339e-02 & 1.902e+04 & 0.355 & 0.722877 & \\
    Position 3 &-1.795e-02&1.259e-02&1.903e+04&-1.426& 0.154005 & \\
    High tone &1.531e-02&4.460e-02&4.358e+03&0.343& 0.731401 & \\
    Low tone &-7.839e-02&3.220e-02&8.169e+03&-2.435& 0.014931 & * \\
    Mid tone &-1.054e-01&4.400e-02&7.421e+03&-2.395& 0.016637 & *  \\
    Rising tone &2.868e-01&6.922e-02&1.022e+04&4.144&3.45e-05 & *** \\
    Breathy:Position 2 &2.104e-02&3.070e-02&1.904e+04&0.685& 0.493043 & \\
    Checked:Position 2 &-3.435e-04&3.598e-02&1.905e+04&-0.010& 0.992383 & \\
    Laryngealized:Position 2 &-6.258e-02&3.227e-02&1.903e+04&-1.940&0.052436 & .  \\
    Breathy:Position 3 &-1.6635e-01&2.901e-02&1.908e+04&5.634& 1.78e-08 & *** \\
    Checked:Position 3 &4.822e-02&3.401e-02&1.909e+04&1.418&0.156192 & \\
    Laryngealized:Position 3 &1.409e-01&3.026e-02&1.906e+04&4.657&3.23e-06 & *** \\
    \lspbottomrule
  \end{tabular}
\end{table}

@@

\begin{table}[!h]
  \centering
  \caption{Results of the statistical model for H1*-A3.}
  \label{tab:H1A3Results}
  \begin{tabular}{lllllll}
    \lsptoprule
     & Estimate & Std. Error & df & t value & p-value & \\ \hline
    \textit{Intercept} & -3.108e-01 & 8.527e-02 & 5.029e+00 & -3.645 & 0.014674 & * \\
    Breathy & 3.870e-01 & 3.951e-02 & 1.014e+04 & 9.795 & < 2e-16 & *** \\
    Checked & -5.464e-01 & 3.694e-02 & 1.274e+04 & -14.790 & < 2e-16 & *** \\
    Laryngealized & -9.881e-02 & 4.886e-02 & 5.909e+03 & -2.022 & 0.043188 & *  \\
    Position 2 & -1.964e-02 & 1.154e-02 & 1.899e+04 & -1.703 & 0.088627 & . \\
    Position 3 & 3.045e-03 & 1.085e-02 & 1.900e+04 & 0.281 & 0.778964 & \\
    High tone & 2.596e-01 & 4.430e-02 & 4.933e+03 & 5.859 & 4.95e-09 & ***  \\
    Low tone & 3.514e-01 & 3.055e-02 & 1.095e+04 & 11.504 & < 2e-16 & *** \\
    Mid tone & 8.087e-02 & 4.212e-02 & 9.234e+03 & 1.920 & 0.054923 & . \\
    Rising tone & 7.312e-01 & 6.457e-02 & 1,431e+04 & 11.323 & < 2e-16 & *** \\
    Breathy:Position 2 & -1.424e-02 & 2.645e-02 & 1.901e+04 & -0.538 & 0.590425 & \\
    Checked:Position 2 & 2.067e-02 & 3.100e-02 & 1.902e+04 & 0.667 & 0.504998 & \\
    Laryngealized:Position 2 & -4.614e-02 & 2.780e-02 & 1.900e+4 & -1.660 & 0.097006 & . \\
    Breathy:Position 3 & -7.101e-02 & 2.500e-02 & 1.904+04 & -2.840 & 0.004516 & *** \\
    Checked:Position 3 & -2.534e-04 & 2.931e-02 & 1.904e+04 & -0.009 & 0.993102 & \\
    Laryngealized:Position 3 & 9.593e-02 & 2.607e-02 & 1.902e+04 & 3.679 & 0.000235 & *** \\
    \lspbottomrule
  \end{tabular}
\end{table}

@@

\begin{table}[!h]
  \centering
  \caption{Results of the statistical model for Residual H1*.}
  \label{tab:H1H2Results}
  \begin{tabular}{lllllll}
    \lsptoprule
     & Estimate & Std. Error & df & t value & p-value & \\ \hline
    \textit{Intercept} & -1.051e-01 & 4.346e-02 & 7.543e+00 & -2.418 & 0.043781 & * \\
    Breathy & 1.502e-01 & 3.184e-02 & 5.952e+03 & 4.716 & 2.46e-06 & *** \\
    Checked & -3.848e-01 & 3.070e-02 & 6.691e+03 & -12.535 & < 2e-16 & *** \\
    Laryngealized & -4.856e-01 & 3.747e-02 & 4.769e+03 & -12.960 & < 2e-16 & *** \\
    Position 2 & 1.369e-02 & 1.027e-02 & 1.902e+04 & 1.333 & 0.182633 & \\
    Position 3 & 4.386e-02 & 9.660e-03 & 1.903e+04 & 4.541 & 5.64e-06 & *** \\
    High tone & 9.129e-02 & 3.345e-02 & 3.138e+03 & 2.729 & 0.006393 & ** \\
    Low tone & 1.854e-01 & 2.441e-02 & 7.861e+03 & 7.595 & 3.44e-14 & *** \\
    Mid tone & 5.386e-02 & 3.332e-02 & 7.052e+03 & 1.616 & 0.106031 & \\
    Rising tone & 3.346e-01 & 5.260e-02 & 9.818e+03 & 6.360 & 2.10e-10 & *** \\
    Breathy:Position 2 & 1.160e-01 & 2.355e-02 & 1.904e+04 & 4.926 & 8.47e-07 & *** \\
    Checked:Position 2 & 1.006e-01 & 2.760e-02 & 1.905e+04 & 3.645 & 0.000268 & *** \\
    Laryngealized:Position 2 & -2.675e-02 & 2.475e-02 & 1.903e+04 & -1.081 & 0.279849 & \\
    Breathy:Position 3 & 1.149e-01 & 2.225e-02 & 1.908e+04 & 5.163 & 2.45e-07 & *** \\
    Checked:Position 3 & -1.033e-01 & 2.608e-02 & 1.909e+04 & -3.959 & 7.55e-05 & *** \\
    Laryngealized:Position 3 & 1.203e-01 & 2.321e-02 & 1.906e+04 & 5.183 & 2.21e-07 & ***\\
    \lspbottomrule
  \end{tabular}
\end{table}
